### 2022-04-05 outputs/paper/paper.Rmd
added relevant files

---
title: "Impact of Internet Use on Canadian Productivity and Recreation"
subtitle: "Brief Overview of Canadian Internet Use Trends"
author: "Haocong Wu"
date: "`r format(Sys.time(), '%d %B %Y')`"
thanks: "Code and data are available at: https://github.com/hcgw0318/2018-Canadian-Internet-Use-Analysis"
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(bookdown)
library(here)
library(forcats)
```

```{r, include=FALSE}
# DATA IMPORT
```

# Introduction

# Data

## Data Source

## Survey Methodology

# Simulation and Modelling

# Results

# Discussion

# References

### 2022-04-06 outputs/paper/paper.Rmd
initial knit check
draft

@@
 ---
-title: "Impact of Internet Use on Canadian Productivity and Recreation"
+title: "Impact of Internet Infrastructure on Canadian Productivity and Recreation"
 subtitle: "Brief Overview of Canadian Internet Use Trends"
@@
 thanks: "Code and data are available at: https://github.com/hcgw0318/2018-Canadian-Internet-Use-Analysis"
-abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
+abstract: "Internet access is critical to many users.
+This paper examines the quality of internet access that users have and how they are impacted by subpar service.
+With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, stability positively impacts user sentiment, whilst speed has diminishing returns past a certain point."
 output:
@@
 
+Keywords: internet use, internet infrastructure, statistics canada, canadian residents, linear regression
+
 # Introduction
 
+Internet access is becoming, if not already considered, a human right.
+It is a standard point of communication used by the majority of the Canadian population.
+The platform that the internet provides can be used to host a variety of activities, ranging from those of productive nature, to others of recreational and leisurely type.
+The growing importance of internet access necessitates a revision of the bare minimums of providing internet service.
+
+This paper examines the causal relations between individual sentiments on their internet use experiences to factors such as connection speed, connection stability, pricing, hours of use, nature of use, and necessity of use.
+
+The following sections of the paper is organized as stated: the Data section discusses the source of the data and its surveying methodology, the Modelling section attempts to formalize a model incorporating various factors to explain the response variable of user sentiment, the Results section discusses the impact of the selected factors, and the Discussion section comments on the results and its validity, potential bias, and weaknesses of the paper.
+
 # Data
+
+[@citeR]
 
@@
 
-# Simulation and Modelling
+# Modelling
 

### 2022-04-27 outputs/eda.Rmd
updated scripts added data content to paper updated raw data dictionary

---
title: "exploratory data analysis"
author: "HC"
date: "21/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(here)
library(forcats)
library(modelsummary)
```

```{r}
cius_data <- read_csv(here("outputs/data/cleaned_data_v2.csv"))
cius_data[] <- lapply(cius_data, factor)
```

frequency plot of province
```{r}
ggplot(cius_data, aes(x=fct_infreq(province)))+
  geom_bar(stat = "count")
```
frequency plot of education
```{r}
ggplot(cius_data, aes(x=fct_infreq(education_level)))+
  geom_bar(stat = "count")
```
frequency plot of age group
```{r}
ggplot(cius_data, aes(x=fct_infreq(age_group)))+
  geom_bar(stat = "count")
```
series of graphs of gender, language prof, employment, immigrant status
```{r}
par(mfrow=c(2,2))
ggplot(cius_data, aes(x=gender))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=official_language_proficient))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=employment_status))+
  geom_bar(stat = "count")
ggplot(cius_data, aes(x=is_immigrant))+
  geom_bar(stat = "count")
```

graph of internet use status
```{r}
ggplot(cius_data, aes(x=uses_internet))+
  geom_bar(stat = "count")
```
frequency plot of hours used
```{r}
ggplot(cius_data, aes(x=hours_used))+
  geom_bar(stat = "count")
```
graph of access at home
```{r}
ggplot(cius_data, aes(x=home_internet_access))+
  geom_bar(stat = "count")
```
summary statistics of household connection type
```{r}
connection_data <- summarise_all(cius_data[,62:73], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "connection_type", values_to = "count")
ggplot(connection_data, aes(x=reorder(connection_type, -count),y=count))+
  geom_bar(stat = "identity")
```

frequency plot of connection speed

```{r}
ggplot(cius_data, aes(x=fct_infreq(connection_speed)))+
  geom_bar(stat = "count")
```

summary statistics of device type
```{r}
device_data <- summarise_all(cius_data[,15:21], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "device_type", values_to = "count")
ggplot(device_data, aes(x=reorder(device_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of communications
```{r}
communications_data <- summarise_all(cius_data[,22:28], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "communications_type", values_to = "count")
ggplot(communications_data, aes(x=reorder(communications_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of information
```{r}
information_data <- summarise_all(cius_data[,30:33], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "information_type", values_to = "count")
ggplot(information_data, aes(x=reorder(information_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of entertainment
```{r}
entertainment_data <- summarise_all(cius_data[,35:43], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "entertainment_type", values_to = "count")
ggplot(entertainment_data, aes(x=reorder(entertainment_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of commerce
```{r}
commerce_data <- summarise_all(cius_data[,45:48], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "commerce_type", values_to = "count")
ggplot(commerce_data, aes(x=reorder(commerce_type, -count),y=count))+
  geom_bar(stat = "identity")
```
summary statistics of other activities
```{r}
other_activities_data <- summarise_all(cius_data[,50:57], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "other_activities_type", values_to = "count")
ggplot(other_activities_data, aes(x=reorder(other_activities_type, -count),y=count))+
  geom_bar(stat = "identity")
```

graph of internet use as source of income

```{r}
ggplot(cius_data, aes(x=internet_income))+
  geom_bar(stat = "count")
```
graph of major or additional source of income

```{r}
income_data <- summarise_all(cius_data[,60:61], table)[-c(1:3),] %>%
  pivot_longer(everything(),names_to = "income_type", values_to = "count")
ggplot(income_data, aes(x=reorder(income_type, -count),y=count))+
  geom_bar(stat = "identity")
```

```{r}
internet_activity <- cbind(select(cius_data,
                                  no_communication,
                                  no_commerce,
                                  no_entertainment,
                                  no_information,
                                  no_other_activities,
                                  internet_income)) %>%
  summarise_all(table)
internet_activity <- internet_activity[-c(2:4),] %>%
  pivot_longer(everything(),names_to = "activity", values_to = "count")
ggplot(internet_activity, aes(x=reorder(activity, -count),y=count))+
  geom_bar(stat = "identity")
```

### 2022-04-27 outputs/paper/paper.Rmd
updated scripts added data content to paper updated raw data dictionary

@@
 ---
-title: "Impact of Internet Infrastructure on Canadian Productivity and Recreation"
+title: "Impact of Internet Infrastructure on Canadian Household Productivity and Recreation"
 subtitle: "Brief Overview of Canadian Internet Use Trends"
@@
 This paper examines the quality of internet access that users have and how they are impacted by subpar service.
-With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, stability positively impacts user sentiment, whilst speed has diminishing returns past a certain point."
+With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, stability positively impacts user household internet usage, whilst speed has diminishing returns past a certain point."
 output:
@@
 ```{r, include=FALSE}
-# DATA IMPORT
+cius_data <- read_csv(here("outputs/data/cleaned_data_v2.csv"))
 ```
@@
 
-This paper examines the causal relations between individual sentiments on their internet use experiences to factors such as connection speed, connection stability, pricing, hours of use, nature of use, and necessity of use.
+This paper examines the causal relations between individual internet usage amounts to factors such as connection speed,nature of use, and necessity of use.
 
-The following sections of the paper is organized as stated: the Data section discusses the source of the data and its surveying methodology, the Modelling section attempts to formalize a model incorporating various factors to explain the response variable of user sentiment, the Results section discusses the impact of the selected factors, and the Discussion section comments on the results and its validity, potential bias, and weaknesses of the paper.
+The following sections of the paper is organized as stated: the Data section discusses the source of the data and its surveying methodology, the Modelling section attempts to formalize a model incorporating various factors to explain the response variable of user household internet use duration, the Results section discusses the impact of the selected factors, and the Discussion section comments on the results and its validity, potential bias, and weaknesses of the paper.
 
@@
 
-[@citeR]
+## Data Source
 
-## Data Source
+The Canadian Internet Use Survey (CIUS) replaced the Household Internet Use Survey (HIUS) in 2005 [@citecius].
+The 2018 of the CIUS aims to record and track the quickly increasing Canadian internet utilization.
+Aspects of internet technologies that Canadians interact with such as online government services, social networks, smartphone use, and online commercial platforms are presented to respondents for their sentiments.
+
+Survey data was provided by the <odesi> portal, maintained by the Ontario Council of University Libraries.
+
+The particular set of data used in this report is a subset of the topics covered by the 2018 CIUS.
 
@@
 
-# Modelling
+Based on the 2018 CIUS documentation [@citecius2018doc], the survey methodology can be summarized by the following.
+
+The CIUS attempts to gauge the population habits regarding various facets of internet use The CIUS' stated target population includes all individuals at and above the age of 15 in Canada, excluding those in the Canadian Territories and excluding institutionalized individuals.
+
+The sampling frame of the GSS relies on the availability of telephone access.
+Households without telephones, as well as households with telephone services not covered by the current frame, were excluded from the sampling frame.
+
+The sampling process consisted of location based random sampling without replacement.
+A target minimum sample size for each province was determined, and this minimum sample size was allocated to the strata within the province.
+
+Each respondent was contacted to complete either an electronic questionnaire, or participate in a computer assisted telephone interview.
+If the respondent refused or was absent, further contact attempts were made through telephone contact until reasonable confirmation of non-response, and the respondent is removed from the pool.
+
+The 2018 CIUS had a target sample size of 15,000. The actual number of respondents was 13,810, the total households contacted was 33,248, and the overall response rate was 41.6%.
+
+## Data Processing Tools and Methodology
+
+Data processing and analysis in this project is done in the R statistical programming language [@citeR]. The following R packages were imported and used for their corresponding purpose:
+
+* `tidyverse` [@citetidyverse] : Used for general logical commands, file importing, and data manipulation
+
+* `knitr` [@citeknitr] : Used to generate and knit the Markdown document to PDF
+
+* `ggplot2` [@citeggplot2] : Used to generate graphs and apply styling
+
+* `kableExtra` [@citekableExtra] : Used to assist in table formatting
+
+* `bookdown` [@citebookdown] : Used to improve the Markdown document formatting
+
+* `here` [@citehere] : Used to simplify file access in a project environment
+
+* `forcats` [@citeforcats] : Used to sort data tables by frequency
+
+## Variables of Interest
+
+```{r}
+ggplot(cius_data, aes(x=fct_infreq(uses_internet), fill=home_internet_access))+
+  geom_bar(stat = "count") +
+  labs(x = "Uses the internet", 
+       y = "Count",
+       title = "Respondents and their usage and access to internet at home",
+       fill = "Home internet access")
+
+```
+
+```{r}
+cius_data$education_level <- cius_data$education_level %>% 
+  factor(levels=c("High school diploma or an equivalency certificate or less",
+                  "Certificate/diploma Univ. below bachelor/College/CEGEP/trade",
+                  "University degree above the bachelor or bachelor's degree",
+                  "Not stated"))
+
+ggplot(cius_data, aes(x=education_level,fill=uses_internet))+
+  geom_bar(stat = "count")+
+  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0))+
+  labs(x = "Education level", 
+       y = "Count",
+       title = "Education level and internet use",
+       fill = "Internet usage")
+```
+
+```{r}
+cius_data$age_group <- cius_data$age_group %>% 
+  factor(levels=c("15 to 24 years",
+                  "25 to 34 years",
+                  "35 to 44 years",
+                  "45 to 54 years",
+                  "55 to 64 years",
+                  "65 years and over"))
+ggplot(cius_data, aes(x=age_group,fill=uses_internet))+
+  geom_bar(stat = "count") + 
+  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
+  labs(x = "Age group", 
+       y = "Count",
+       title = "Age group and internet use",
+       fill = "Internet usage")
+```
+
+```{r}
+connection_data <- summarise_all(cius_data[,62:73], table)[-c(1:3),] %>%
+  pivot_longer(everything(),names_to = "connection_type", values_to = "count")
+ggplot(connection_data, aes(x=reorder(connection_type, -count),y=count))+
+  geom_bar(stat = "identity") + 
+  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0)) + 
+  labs(x = "Connection type", 
+       y = "Count",
+       title = "Home internet connection type")
+```
+```{r}
+internet_activity <- cbind(select(cius_data,
+                                  no_communication,
+                                  no_commerce,
+                                  no_entertainment,
+                                  no_information,
+                                  no_other_activities,
+                                  internet_income)) %>%
+  summarise_all(table)
+internet_activity <- internet_activity[-c(2:4),] %>%
+  rename(Communication=no_communication,
+         Commerce=no_commerce,
+         Entertainment=no_entertainment,
+         Information=no_information,
+         Other=no_other_activities,
+         Income=internet_income) %>%
+  pivot_longer(everything(),names_to = "activity", values_to = "count")
+ggplot(internet_activity, aes(x=reorder(activity, -count),y=count))+
+  geom_bar(stat = "identity") + 
+  labs(x = "Internet usage type", 
+       y = "Count",
+       title = "Purposes of internet usage")
+```
+
+```{r}
+cius_data$hours_used <- cius_data$hours_used %>% 
+  factor(levels=c("Less than 5 hours per week",
+                  "5 to less than 10 hours per week",
+                  "10 to less than 20 hours per week",
+                  "20 to less than 40 hours per week",
+                  "40 hours or more per week",
+                  "Not stated",
+                  "Valid skip"))
+ggplot(cius_data, aes(x=hours_used))+
+  geom_bar(stat = "count") + 
+  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0)) + 
+  labs(x = "Duration of use",
+       y = "Count",
+       title = "Hours of internet use per week (excluding entertainment)")
+```
 
 # Results
+
+```{r}
+set.seed(318)
+regression_dat <- cius_data[cius_data$uses_internet=="Yes" &
+                              cius_data$home_internet_access=="Yes" &
+                              !(cius_data$hours_used=="Valid skip" | cius_data$hours_used=="Not stated"),]
+regression_dat <- regression_dat %>% 
+  mutate(hours_used=case_when(hours_used=="Less than 5 hours per week"~
+                                sample(x = c(0:4), size = n(),replace = TRUE),
+                              hours_used=="5 to less than 10 hours per week"~
+                                sample(x = c(5:9), size = n(),replace = TRUE),
+                              hours_used=="10 to less than 20 hours per week"~
+                                sample(x = c(10:19), size = n(),replace = TRUE),
+                              hours_used=="20 to less than 40 hours per week"~
+                                sample(x = c(20:39), size = n(),replace = TRUE),
+                              hours_used=="40 hours or more per week"~
+                                sample(x = c(40:60), size = n(),replace = TRUE)))
+```
+
+```{r}
+lm_model <- lm(data = regression_dat, hours_used ~ 
+                 gender + 
+                 education_level + 
+                 age_group)
+modelsummary(lm_model)
+summary(lm_model)
+```
 

### 2022-04-27 scripts/00_simulation.R

@@ -1,0 +1,20 @@
+library(tidyverse)
+
+set.seed(318)
+population <- 13810
+sim_dataset <- tibble(age_group = sample(x = c(1:6),
+                                         size = population,
+                                         replace = TRUE),
+                      gender = sample(x = c(1:2),
+                                      size = population,
+                                      replace = TRUE))
+
+sim_dataset <- sim_dataset %>% 
+  mutate(age_group=case_when(age_group==1~"15 to 24 years",
+                             age_group==2~"25 to 34 years",
+                             age_group==3~"35 to 44 years",
+                             age_group==4~"45 to 54 years",
+                             age_group==5~"55 to 64 years",
+                             age_group==6~"65 years and over",),
+         gender=case_when(gender==1~'Male',
+                          gender==2~'Female'))

### 2022-04-27 scripts/01_clean_data_values.R

@@ -1,0 +1,56 @@
+library(janitor)
+library(tidyverse)
+library(here)
+
+raw_data <- read_csv(here("inputs/data/raw_data.csv"))[,c(2:425)] %>%
+  mutate(AGE_GRP = as.numeric(AGE_GRP))
+labels_raw <- read_file(here("inputs/data/response_labels.txt"))
+
+labels_raw_tibble <- as_tibble(str_split(labels_raw, "\"\r\n")[[1]]) %>% 
+  mutate(value = str_remove(value, "label define   ")) %>%
+  mutate(value = str_remove(value, "_FMT")) %>%
+  mutate(value = str_remove_all(value, " ///")) %>%
+  mutate(value = str_replace(value, "[ ]{2,}", "XXX")) %>% 
+  mutate(splits = str_split(value, "XXX")) %>% 
+  rowwise() %>% 
+  mutate(variable_name = splits[1], cases = splits[2]) %>% 
+  mutate(cases = str_replace_all(cases, "\n [ ]{2,}", "")) %>%
+  select(variable_name, cases) %>% 
+  drop_na()
+
+labels_raw_tibble <- labels_raw_tibble %>% 
+  mutate(splits = str_split(cases, "[ ]{0,}\"[ ]{0,}"))
+
+add_cw_text <- function(x, y){
+  if(!is.na(as.numeric(x))){
+    x_new <- paste0(y, "==", x,"~")
+  }
+  else{
+    x_new <- paste0("\"",x,"\",")
+  }
+  return(x_new)
+}
+
+cw_statements <- labels_raw_tibble %>% 
+  rowwise() %>% 
+  mutate(splits_with_cw_text = list(modify(splits, add_cw_text, y = variable_name))) %>% 
+  mutate(cw_statement = paste(splits_with_cw_text, collapse = "")) %>% 
+  mutate(cw_statement = paste0("case_when(", cw_statement,"TRUE~\"NA\")")) %>% 
+  mutate(cw_statement = str_replace(cw_statement, ",\"\",",",")) %>% 
+  select(variable_name, cw_statement)
+
+cw_statements <- 
+  cw_statements %>% 
+  mutate(variable_name = str_remove_all(variable_name, "\r")) %>%
+  mutate(variable_name = str_remove_all(variable_name, "\n")) %>%
+  mutate(cw_statement = str_remove_all(cw_statement, "\r")) %>%
[…]

### 2022-04-27 scripts/02_clean_data_names.R

@@ -1,0 +1,178 @@
+library(janitor)
+library(tidyverse)
+library(here)
+
+cius_data <- read_csv(here("outputs/data/cleaned_data_v1.csv"))
+
+
+#demography
+
+demo_dat <- cius_data %>% select(
+  PROVINCE,
+  G_EDUC,
+  GENDER,
+  AGE_GRP,
+  LAN_Q01,
+  EMP,
+  BPR_Q16,
+  VS_USER,
+  GV_USER,
+  SN_USER,
+  SM_USER
+)%>% rename(province=PROVINCE,
+            education_level=G_EDUC,
+            gender=GENDER,
+            age_group=AGE_GRP,
+            official_language_proficient=LAN_Q01,
+            employment_status=EMP,
+            is_immigrant=BPR_Q16,
+            video_streaming_use=VS_USER,
+            gov_serv_use=GV_USER,
+            social_ntwk_use=SN_USER,
+            smartphone_use=SM_USER)
+
+#internet access
+
+access_dat <- cius_data %>% select(
+  AC_160A, #used internet
+  AC_180A, #access at home
+)%>% rename(uses_internet=AC_160A,
+            home_internet_access=AC_180A)
+
+#internet use
+
+usage_dat <- cius_data %>% select(
+  UI_210A, #hours used
+  UI_220A, #device type smartphone
+  UI_220B, #device type laptop
+  UI_220C, #device type tablet
+  UI_220D, #device type other mobile device
+  UI_220E, #device type desktop
+  UI_220F, #device type media streaming device
+  UI_220G, #device type other internet device
+  UI_241A, #communication type email
+  UI_241B, #communication type im
+  UI_241C, #communication type social network
+  UI_241D, #communication type voice/video call
+  UI_241E, #communication type dating svc
+  UI_241F, #communication type content publishing
+  UI_241G, #communication type blog
+  UI_241H, #communication type none
+  UI_242A, #information type news
+  UI_242B, #information type weather
+  UI_242C, #information type directions/maps
+  UI_242D, #information type research
+  UI_242E, #information type none
+  UI_243A, #entertainment type music
+  UI_243B, #entertainment type podcasts
+  UI_243C, #entertainment type streaming svc
+  UI_243D, #entertainment type video-sharing sites
+  UI_243E, #entertainment type sports subscription
+  UI_243F, #entertainment type television stream
+  UI_243G, #entertainment type books/magazines
+  UI_243H, #entertainment type video games
[…]

### 2022-04-27 outputs/paper/paper.Rmd
added results content

@@
 This paper examines the quality of internet access that users have and how they are impacted by subpar service.
-With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, stability positively impacts user household internet usage, whilst speed has diminishing returns past a certain point."
+With data collected by Statistics Canada in the Canadian Internet Use Survey, and modelled with linear regression, age has a negative correlation with user household internet usage, whilst education level has a strong positive correlation."
 output:
@@
 library(forcats)
+library(modelsummary)
 ```
@@
 
-## Variables of Interest
+* `modelsummary` [@citemodelsummary] : Used to display regression model information
 
-```{r}
-ggplot(cius_data, aes(x=fct_infreq(uses_internet), fill=home_internet_access))+
-  geom_bar(stat = "count") +
-  labs(x = "Uses the internet", 
-       y = "Count",
-       title = "Respondents and their usage and access to internet at home",
-       fill = "Home internet access")
+## Relevant variables
 
+Variables such as whether the respondent used the internet and had an internet connection at home were used to identify the subset of the sample whose internet usage time is the response of interest.
+Variables likely to explain variations in internet usage time contained in the survey were age, education level, and purpose of internet usage.
+
+```{r, echo=FALSE}
+demo_dat <- cius_data %>% select(age_group, gender) %>%
+  group_by(age_group, gender) %>%
+  tally() %>%
+  pivot_wider(names_from = gender, values_from = n) %>% 
+  remove_rownames %>% 
+  column_to_rownames(var="age_group")
+
+prov_dat <- cius_data %>% select(province) %>% group_by(province) %>% tally()
+colnames(prov_dat) <- c("Province", "Count")
 ```
 
-```{r}
-cius_data$education_level <- cius_data$education_level %>% 
-  factor(levels=c("High school diploma or an equivalency certificate or less",
-                  "Certificate/diploma Univ. below bachelor/College/CEGEP/trade",
-                  "University degree above the bachelor or bachelor's degree",
-                  "Not stated"))
-
-ggplot(cius_data, aes(x=education_level,fill=uses_internet))+
-  geom_bar(stat = "count")+
-  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0))+
-  labs(x = "Education level", 
-       y = "Count",
-       title = "Education level and internet use",
-       fill = "Internet usage")
+```{r demotab1, fig.cap="Survey Respondent Demographics - Location", echo=FALSE}
+# table
+kbl(prov_dat, caption = "Location Demographics", booktabs = T, linesep = "") %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), latex_options = "HOLD_position") %>%
+  kable_paper("striped", full_width = F)
 ```
 
-```{r}
-cius_data$age_group <- cius_data$age_group %>% 
-  factor(levels=c("15 to 24 years",
-                  "25 to 34 years",
-                  "35 to 44 years",
-                  "45 to 54 years",
-                  "55 to 64 years",
-                  "65 years and over"))
-ggplot(cius_data, aes(x=age_group,fill=uses_internet))+
-  geom_bar(stat = "count") + 
-  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
-  labs(x = "Age group", 
-       y = "Count",
-       title = "Age group and internet use",
-       fill = "Internet usage")
+```{r demotab2, fig.cap="Survey Respondent Demographics - Age and Gender", echo=FALSE}
+# table
+kbl(demo_dat, caption = "Age and Gender Demographics", booktabs = T, linesep = "") %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), latex_options = "HOLD_position") %>%
+  kable_paper("striped", full_width = F)
 ```
 
-```{r}
+In Table \@ref(tab:demotab1) and Table \@ref(tab:demotab2), the survey demographics are shown.
+It can be seen that the sample population is skewed towards more female respondents, and skewed towards older respondents.
+
+```{r, echo=FALSE}
 connection_data <- summarise_all(cius_data[,62:73], table)[-c(1:3),] %>%
   pivot_longer(everything(),names_to = "connection_type", values_to = "count")
+```
+
+```{r homeconnection, fig.cap="Respondent home connection type", echo=FALSE}
 ggplot(connection_data, aes(x=reorder(connection_type, -count),y=count))+
   geom_bar(stat = "identity") + 
+  scale_y_continuous() +
   theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0)) + 
@@
 ```
-```{r}
+
+In Figure \@ref(fig:homeconnection), the three most popular types of home internet connections are cable, fiber optics, and mobile data.
+After that, approximately 7% of respondents are not sure what type of connection they have.
+
+```{r, echo=FALSE}
 internet_activity <- cbind(select(cius_data,
@@
   pivot_longer(everything(),names_to = "activity", values_to = "count")
+```
+
+```{r usepurpose, fig.cap="Respondent internet use purpose", echo=FALSE}
 ggplot(internet_activity, aes(x=reorder(activity, -count),y=count))+
@@
 
-```{r}
+In Figure \@ref(fig:usepurpose), common uses of the internet are displayed.
+
+```{r, echo=FALSE}
 cius_data$hours_used <- cius_data$hours_used %>% 
@@
                   "Valid skip"))
+```
+
+```{r internethours, fig.cap="Respondent weekly internet use", echo=FALSE}
 ggplot(cius_data, aes(x=hours_used))+
@@
 
+In Figure \@ref(fig:internethours), the distribution of weekly internet usage of the respondents is shown. The most common amount of time spent on the internet per week is between 5 to 10 hours.
+
 # Results
 
-```{r}
+```{r internetuse, fig.cap="Respondent internet use and home internet access", echo=FALSE}
+#graph
+ggplot(cius_data, aes(x=fct_infreq(uses_internet), fill=home_internet_access))+
+  geom_bar(stat = "count") +
+  labs(x = "Uses the internet", 
+       y = "Count",
+       title = "Respondents and their usage and access to internet at home",
+       fill = "Home internet access")
+```
+
+In Figure \@ref(fig:internetuse), it is shown that most users use the internet in some way. The majority of respondents that do not use the internet also do not have an internet connection at home.
+
+```{r, echo=FALSE}
+cius_data$education_level <- cius_data$education_level %>% 
+  factor(levels=c("High school diploma or an equivalency certificate or less",
+                  "Certificate/diploma Univ. below bachelor/College/CEGEP/trade",
+                  "University degree above the bachelor or bachelor's degree",
+                  "Not stated"))
+
+```
+
+```{r educationlevel, fig.cap="Respondent education level and internet use", echo=FALSE}
+ggplot(cius_data, aes(x=education_level,fill=uses_internet))+
+  geom_bar(stat = "count")+
+  theme(axis.text.x = element_text(angle = -15, vjust = 1, hjust=0))+
+  labs(x = "Education level", 
+       y = "Count",
+       title = "Education level and internet use",
+       fill = "Internet usage")
+```
+
+In Figure \@ref(fig:educationlevel), it can be seen that internet usage proportion increases with respondent education level.
+
+```{r, echo=FALSE}
+cius_data$age_group <- cius_data$age_group %>% 
+  factor(levels=c("15 to 24 years",
+                  "25 to 34 years",
+                  "35 to 44 years",
+                  "45 to 54 years",
+                  "55 to 64 years",
+                  "65 years and over"))
+```
+
+```{r agegroup, fig.cap="Respondent age group and internet use", echo=FALSE}
+ggplot(cius_data, aes(x=age_group,fill=uses_internet))+
+  geom_bar(stat = "count") + 
+  scale_x_discrete(guide = guide_axis(n.dodge = 2)) + 
+  labs(x = "Age group", 
+       y = "Count",
+       title = "Age group and internet use",
+       fill = "Internet usage")
+```
+
+In Figure \@ref(fig:agegroup), as age group increases, the proportion of the group that does not use the internet increases.
+
+```{r, echo=FALSE, warning=FALSE}
 set.seed(318)
@@
                                 sample(x = c(40:60), size = n(),replace = TRUE)))
-```
+regression_dat <- regression_dat %>%
+  mutate_at(c("no_communication",
+              "no_commerce",
+              "no_entertainment",
+              "no_information",
+              "no_other_activities",
+              "internet_income"), funs(case_when(.=="Yes"~TRUE,
+                                                 TRUE~FALSE)))
+regression_dat <- regression_dat %>%
+  mutate_at(vars(62:73), funs(case_when(.=="Yes"~TRUE,
+                                        TRUE~FALSE)))
+regression_dat$connection_type <- names(regression_dat)[62:73][max.col(regression_dat[62:73], 'first')]
 
-```{r}
 lm_model <- lm(data = regression_dat, hours_used ~ 
@@
                  education_level + 
-                 age_group)
-modelsummary(lm_model)
-summary(lm_model)
+                 age_group +
+                 no_communication +
+                 no_commerce +
+                 no_entertainment +
+                 no_information +
+                 no_other_activities +
+                 internet_income)
 ```
+
+```{r regressionmodel, fig.cap="2018 Canadian Internet Use Regression Model", echo=FALSE}
+modelsummary(lm_model, fmt=1, title = "Model of weekly internet use quantity")
+```
+
+In Table \@ref(tab:regressionmodel), the regression model of internet use hours as a function of gender, education level, age group, and internet activities is displayed.
 
@@
 
+## Greater proportion of female respondents
+
+## Greater proportion of older respondents
+
+## Education and age trends
+
+## Weaknesses and improvements
+
 # References

### 2022-04-28 outputs/paper/paper.Rmd
finalize paper

@@
 ---
-title: "Impact of Internet Infrastructure on Canadian Household Productivity and Recreation"
+title: "Canadian Household Internet Productivity and Recreation Trends 2018"
 subtitle: "Brief Overview of Canadian Internet Use Trends"
@@
 
-This paper examines the causal relations between individual internet usage amounts to factors such as connection speed,nature of use, and necessity of use.
+This paper examines the causal relations between individual internet usage amounts to factors such as connection speed, nature of use, and necessity of use.
 
@@
 
+\newpage
+
 ```{r, echo=FALSE}
@@
 After that, approximately 7% of respondents are not sure what type of connection they have.
+
+\newpage
 
@@
        y = "Count",
-       title = "Hours of internet use per week (excluding entertainment)")
+       title = "Hours of internet use per week")
 ```
@@
 In Figure \@ref(fig:internethours), the distribution of weekly internet usage of the respondents is shown. The most common amount of time spent on the internet per week is between 5 to 10 hours.
+
+\newpage
 
@@
 In Figure \@ref(fig:internetuse), it is shown that most users use the internet in some way. The majority of respondents that do not use the internet also do not have an internet connection at home.
+
+\newpage
 
@@
 In Figure \@ref(fig:agegroup), as age group increases, the proportion of the group that does not use the internet increases.
+
+\newpage
 
@@
 ```{r regressionmodel, fig.cap="2018 Canadian Internet Use Regression Model", echo=FALSE}
-modelsummary(lm_model, fmt=1, title = "Model of weekly internet use quantity")
+modelsummary(lm_model, fmt=1, title = "Model of weekly internet use quantity") %>%
+  kable_styling(bootstrap_options = c("hover", "condensed"), latex_options = "HOLD_position") %>%
+  kable_paper("striped", full_width = F)
 ```
@@
 
+From data displayed in Table \@ref(tab:demotab1) and graphed, the respondent demographic can be seen to have a larger female representation.
+This observation could be caused by a larger proportion of Canadian females being available to pick up the phone due to childcare and household work.
+
 ## Greater proportion of older respondents
+
+Despite the decreasing proportion of individuals at each age group that uses the internet shown in Figure \@ref(fig:agegroup), the number of respondents increases with each older age group.
+This at first glance seems conflicting with the fact that the CIUS is  conducted primarily on an online platform.
+A reason for this occurrence may be that the older age groups have more leisure time, and are willing to use their leisure time to respond to this survey.
 
@@
 
+As education level increases, the proportion of respondents that use the internet increases, as shown in Figure \@ref(fig:educationlevel).
+In addition, the amount of time that one uses the internet per week increases with education level as well.
+This result likely arises due to the utility if not necessity of utilizing online resources to progress with further education, as well as the fact that individuals who undertake further education would be more inclined to adapt and utilize new technology.
+
+The converse is true with age, the younger the respondent, the more likely they are to both use the internet and use it for a longer duration per week, this is depicted by Figure \@ref(fig:agegroup).
+This occurrence is explainable by the attraction of online entertainment of social media, video games, and media consumption platforms to younger audiences.
+The younger the individual, the more accustomed they are to seek entertainment through online means.
+
 ## Weaknesses and improvements
 
+This paper attempts to address the current populations internet usage tendencies, and focuses on the duration of internet use.
+The respondents' demographics, connection type, and usage purposes were variables that were analyzed.
+Given that the goal of the paper is to find correlation between these factors and internet usage hours, the analysis that consisted of the factors of gender, age, education, and usage nature may be lacking.
+The analysis can be supplemented with the consideration of factors such as household economy, urban and rural living distinctions, and whether anyone in their household utilized the internet.
+These factors were not included in the survey, and can be collected from a follow-up supplemental survey.
+
+Another weakness of the analysis is making the assumption of generating uniform random ages for each age group to perform linear regression.
+This assumption was made due to the lack of more granular internet usage data.
+To improve on this aspect, perhaps ordinal regression may be a better fit to generate a model.
+
+\newpage
+
 # References

### 2022-04-28 scripts/00_simulation.R

@@ -7,6 +7,9 @@ sim_dataset <- tibble(age_group = sample(x = c(1:6),
                                          replace = TRUE),
                       gender = sample(x = c(1:2),
                                       size = population,
+                                      replace = TRUE),
+                      education = sample(x = c(1:3),
+                                      size = population,
                                       replace = TRUE))
 
 sim_dataset <- sim_dataset %>% 
@@ -17,4 +20,11 @@ sim_dataset <- sim_dataset %>%
                              age_group==5~"55 to 64 years",
                              age_group==6~"65 years and over",),
          gender=case_when(gender==1~'Male',
-                          gender==2~'Female'))
+                          gender==2~'Female'),
+         education=case_when(education==1~"High school diploma or an equivalency certificate or less",
+                             education==2~"Certificate/diploma Univ. below bachelor/College/CEGEP/trade",
+                             education==3~"University degree above the bachelor or bachelor's degree"))
+
+ggplot(sim_dataset, aes(x=gender, fill=age_group)) + geom_bar(stat = 'count')
+ggplot(sim_dataset, aes(x=gender, fill=education)) + geom_bar(stat = 'count')
+
